```{r echo=FALSE}
yml_content <- yaml::read_yaml("chapterauthors.yml")
author <- yml_content[["whatIsGeomatics"]][["author"]]
```
# What is Geomatics? {#what-is-geomatics}

Written by
```{r results='asis', echo=FALSE}
cat(author)
```

We encounter and use geographic information on a regular basis in our everyday lives. Whether it is finding directions to a retailer that has an item in stock you want to buy or recording the path of your last morning jog, you have probably used a Geographic Information System (GIS) and not even realized it. 

  :::: {.box-content .learning-objectives-content}
  
  ::: {.box-title .learning-objectives-top}
  ## Learning Objectives {-}
  ::: 
  1. 
  ::::
  
  ## Key Terms {-}
  
  Geomatics, Geographic Information Systems, Remote Sensing

## The Science and Technology of Geomatics

**Geomatics** is the science and technology of collecting geographic data and converting it to geographic information for use in a wide variety of industries. As a technical field, it encompasses many different work processes including surveying, remote sensing, global navigation satellite systems, geospatial analysis, and information technology and systems management. In turn, these processes support a wide variety of spatial decision-making such as urban planning, ecological conservation, forest management, real-time planetary systems monitoring, and rapid response to natural disasters. Many more emerging technologies such as self-driving vehicles, ride-sharing apps, and augmented reality video games depend directly on the science and technology of geomatics.

## A Brief History of Geomatics in Canada

remote sensing is any method of gathering information about an object, or objects, without physical contact.

## Information Systems

An **information system** is used to store, code, and recall information. In the Information Age, we are surrounded and depend heavily on information systems such as financial systems that record the transactions in your bank account or navigation systems that tell you the fastest route to a destination or autonomous vehicles such as the SkyTrain rapid transit network in Vancouver, Canada that moves more than a half million people every day across the Metro Vancouver region. These are all examples of systems that require high synchronization and integration of many varied sources of information in order to move people and assets around. It should come as no surprise then that information systems and information technology contribute significantly to nearly every sector of developed and developing economies. 

What makes a **Geographic Information System** different from other information systems is the type of information that is handled: geographic location. With a GIS you can know the quantity and quality of something and the __location__ of that event, activity, or feature. For example, you might have recorded your heart rate and timed your morning jog using your phone or other fitness tracking device. Since you also have the location or coordinates of your jog, you can calculate or derive additional information such as speed (distance per time), the total distance you jogged, and also your elevation above sea level. You could even relate your heart rate to different locations along your jog to better understand your performance on different terrain. This is an example of a GIS at work. You are storing geographic information into a system that allows you to code the information with different qualities (e.g., the type of surface your ran on) and quantities (e.g., your heart rate) and then recall that information in a way that allows you to explore trends and ask and answer __spatial__ questions.

